Optifcal Observations of an Extraordinarily Young Type Ia Supernova iPTF16abc

### 2017-01-30 doc/ms.tex
backup
Include changes to abstract and introduction

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section defines new commands for comments from co-authors
%%
\newcommand{\ycao}[1]{{\color{red} ycao: {#1}}}
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

  We report discovery of an extraordinarily young Type Ia supernova
  iPTF16abc by the intermediate Palomar Transient Factory. Our first
  observation was made only $0.18$ days after the onset of the
  supernova explosion. Our analysis shows that the early emission of
  this supernova dominantly comes from the radioactive decay instead
  of supernova shock breakout or collision between the ejecta and a
  possible companion star. The fast initial rise of the supernova, the
  velocity evolution of spectral lines, and the strong and short-lived
  carbon feature in the earliest spectra together provide strong
  evidence of strong mixing of synthesized radioactive elements during
  the supernova explosion.

@@

Although Type Ia supernovae (SNe Ia) have been extensively used as
standardizable candles, their progenitor scenarios and explosion
physics are still in debate (see a recent review by
\citealt{2014ARA&A..52..107M}). Detailed extermely early-phase
observations are one of the most promising avenues to further
constrain this problem.

While the shock breakout of a SN Ia occurs on a sub-second timescale,
the subsequent quasi-adiabatic expanding and cooling of the unbinded
ejecta produces thermal emissions that can be used to infer the
original size of the exploding star
\citep{2010ApJ...708..598P,2011ApJ...728...63R}. Comparing models of
these cooling emissions to the earliest-phase data of SN2011fe,
\citet{2012ApJ...744L..17B} concluded that the radius of the
progenitor star is $\lesssim0.01\sr$ where $\sr$ is the solar
radius. Combining this size constraint and the measured ejecta mass to
derive the mean density of the progenitor star, we confirmed that the
progenitor star is compact and degenerate. Admittedly, due to the
initial small surface area of the progenitor star, the shock cooling
emission of a SN Ia decays drastically as the ejecta expands. Given
typical parameters of a SN Ia, this thermal emission is visible from
events up to $\sim 10\,\textrm{Mpc}$ within one day of their explosions.

Another expectation from the extremely early-phase observations of a
SN Ia is the excess emission from collisions between SN ejecta and a
companion star, a natural consequence from the single-degenerate
progenitor hypothesis \citep{1973ApJ...186.1007W,2010ApJ...708.1025K}.
In a low-velocity SN Ia iPTF14atg, \citet{2015Natur.521..328C} for the
first time detected a strong and declining ultraviolet pulse within a
few days of the SN explosion which is best interpreted as the
SN-companion collision. This signature has been searched in a number
of nearby, early-phase normal SNe Ia, but most of these studies result
in no detection
\citep{2010ApJ...722.1691H,2011ApJ...741...20B,2012ApJ...744...38F,
  2012ApJ...744L..17B,2015Natur.521..332O,
  2013ApJ...778L..15Z,2015ApJ...799..106G,2016ApJ...826..144S,
  2015ApJS..221...22I}. The exception is SN2012cg which
\citet{2016ApJ...820...92M} claimed detection of blue excess in the
earliest-phase light curve and attribute it to SN-companion collision.
However, this statement is recently challenged by
\citet{2016arXiv161007601S}. In fact, it is not surprising that no
SN-companion collision has been observationally confirmed, because
only up to $\sim 10\%$ of events from the single-degenerate channel
have the preferred binary geometry for us to see the collision
signatures.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section defines new commands for comments from co-authors
%%
\newcommand{\ycao}[1]{{\color{red} ycao: {#1}}}
\newcommand{\amiller}[1]{{\color{blue} amiller: {#1}}}
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

  We report the discovery of an extraordinarily young Type Ia supernova,
  iPTF16abc, by the intermediate Palomar Transient Factory. Our first
  observation was made only $0.18$~d after the onset of the
  supernova explosion. Our analysis shows that the early emission from iPTF16abc is dominated by radioactive decay, and that contributions from supernova shock breakout or the ejecta colliding with a stellar companion are minimal. The combination of (i) a fast initial rise, (ii) the
  early exponential decline in the photospheric velocity, and (iii) the significant presence of ionized carbon in the earliest spectra together provide
  evidence for the strong mixing of radioactive elements in the supernova ejecta. \amiller{There should be something about dark times/lack of dark time for this SN in the abstract.}

@@

Although Type Ia supernovae (SNe Ia) have been extensively used as
standardizable candles, their progenitors and explosion
physics are still debated (see a recent review by
\citealt{2014ARA&A..52..107M}). Extremely detailed
observations in the hours to days after explosion are one of the most promising avenues to further
constrain this problem.

While the shock breakout of a SN Ia occurs on a sub-second timescale,
the subsequent quasi-adiabatic expanding and cooling of the unbound
ejecta produces thermal emission that can be used to infer the
original size of the exploding star
\citep{2010ApJ...708..598P,2011ApJ...728...63R}. Comparing models of
this cooling emission to the earliest-phase data of SN~2011fe,
\citet{2012ApJ...744L..17B} concluded that the radius of the SN~2011fe progenitor is $\lesssim0.01\sr$, where $\sr$ is the solar
radius. Combining this size constraint and the measured ejecta mass to
derive the mean density of the progenitor star, we confirmed that the
progenitor star is compact and degenerate\ycao{Ref?}. \ycao{a transition is missing here - not sure how the second half of this paragraph relates to the first} Admittedly, due to the
initial small surface area of the progenitor star, the shock cooling
emission of a SN Ia decays drastically as the ejecta expands. Given
typical parameters for a SN Ia, this thermal emission is visible from
events up to $\sim 10\,\textrm{Mpc}$ within one day of their explosions.

SN Ia resulting from a white dwarf and non-degenerate companion, referred to as the single-degenerate progenitor, early-phase observations may detect excess emission due to the collision of the SN ejecta and the non-degenerate companion \citep{1973ApJ...186.1007W,2010ApJ...708.1025K}. This excess emission was first detected in iPTF14atg,  \citep{2015Natur.521..328C}, a low-velocity SN Ia with a significant and declining ultraviolet pulse detected within a
few days of the SN explosion. This signature is best interpreted as a
SN ejecta-companion collision. Many studies have searched for this excess emission, with most resulting in non-detections
\citep{2010ApJ...722.1691H,2011ApJ...741...20B,2012ApJ...744...38F,
  2012ApJ...744L..17B,2015Natur.521..332O,
  2013ApJ...778L..15Z,2015ApJ...799..106G,2016ApJ...826..144S,
  2015ApJS..221...22I}. An exception is SN2012cg, 
  as a blue excess in its early-phase light curve is attributed to an ejecta-companion collision in \citet{2016ApJ...820...92M} (though see \citealt{2016arXiv161007601S} for an alternative interpretation). 
With only $\sim 10\%$ of single-degenerate progenitors occupying the preferred binary geometry for us to see the collision
signatures, it is not surprising that there are few confirmed detections in the literature \ycao{ref?}.
